---
title: "TitanicR"
author: "Michael Doherty"
date: "26/10/2020"
output: html_document
---

#### REQUIRED PACKAGES ####
```{r}
library(caret)
library(naniar)
library(ggplot2)
library(tidyverse)
library(ggcorrplot)
library(randomForest)
library(stringr)
library(outliers)
library(EnvStats)
```

#### CLEANING ####

# IMPORT DATA
```{r}
train <- read.csv('train.csv')
```

# RE-ORDERING TARGET COLUMN
```{r}
train <- train[c(1,3:12,2)]
```

# SETTING SURVIVED VARIABLE TO EITHER YES/NO
```{r}
train$Survived[train$Survived > 0] <- "YES"
train$Survived[train$Survived == 0] <- "NO"
train$Survived <- as.factor(train$Survived)
```

#### OUTLIER DETECTION ####
```{r}
rosnerTest(train$Age, k=3)
rosnerTest(train$SibSp, k=50)
rosnerTest(train$Parch, k=50)
rosnerTest(train$Fare, k=60)
```



#### FEATURE ENGINEERING ####

# family size

# Extract Title from name
```{r}
train$title <- str_split_fixed(train$Name, " ", 3)[,2]
```


# Extract the numeric from Ticket
```{r}
train$numTicket <- gsub("[^0-9\\]", "", train$Ticket)
```

# Extracting the first alphabetic character from the ticket
```{r}
train$char1Ticket <- substr(trimws(gsub('[[:digit:]]+', '', train$Ticket)), 1, 1)
```

# Extracting the first alphabetic character from the cabin
```{r}
train$charCabin <- substr(trimws(gsub('[[:digit:]]+', '', train$Cabin)), 1, 1)
```

# Extracting the mean of the cabin numerics
```{r}
train$meanCabin <- unlist(lapply(lapply(strsplit(gsub("[^0-9\\ ]", "", train$Cabin), " "), as.numeric), mean))
```

# Extracting the number of cabins for each ticket
```{r}
train$nCabins <- lengths(strsplit(as.character(train$Cabin), " "))
```


# FIXING DATA TYPES AND MAKING NAs CONSISTENT
```{r}
NAs <- c("?", "none", "NA", "NaN", "null", "N A", "N/A", "Null", "None", " ", "")
train <- type.convert(train, na.strings = NAs)
```

# RE-ORDERING AGAIN
```{r}
train
train <- train[c(1:11,13:(ncol(train)),12)]
```

# DROPPING REDUNDANT COLUMNS
```{r}
train <- train[c(1,2,4:7,9,11:18)]
train
```

#### EDA ####

# How many NAs?
```{r}
vis_miss(train)
sapply(train, function(x) sum(is.na(x)))
```

# Frequency distibution of Survival
```{r}
train %>%
  ggplot(aes(x= Survived)) +
  geom_bar(width = 0.5, fill='steelblue') +
  ylim(0, 600) +
  geom_text(stat = 'count', aes(label=..count..), vjust=-1)+
  ggtitle('Survival Distribution') +
  xlab('Survival Outcome')
```

# Passenger Class Distribution
```{r}
train %>%
  ggplot(aes(x= Pclass)) +
  geom_bar(width = 0.5, fill='steelblue') +
  ylim(0, 600) +
  geom_text(stat = 'count', aes(label=..count..), vjust=-1)+
  ggtitle('Class Distribution')+
  xlab('Passenger Class')
```


# Passenger Sex Distribution
```{r}
train %>%
  ggplot(aes(x= Sex)) +
  geom_bar(width = 0.5, fill='steelblue') +
  ylim(0, 600) +
  geom_text(stat = 'count', aes(label=..count..), vjust=-1)+
  ggtitle('Sex Distribution')+
  xlab('Passenger Sex')
```

# Passenger Age Distribution
```{r}
train %>%
  ggplot(aes(x= Age)) +
  geom_density(color='darkblue', fill='lightblue') +
  ggtitle('Passenger Age Density Plot') +
  xlab('Passenger Age (years)')
```

# Sibling Spouse Distribution
```{r}
train %>%
  ggplot(aes(x= SibSp)) +
  geom_bar(width = 0.5, fill='steelblue') +
  ylim(0, 250) +
  geom_text(stat = 'count', aes(label=..count..), vjust=-1)+
  ggtitle('Sibling/Spouse Distribution')+
  xlab('# Of siblings or spouses onboard')
```

# Mixed data-type correlation plot
```{r}
source('cor2 correlations.r')
cor2(train)
ggcorrplot(cor2(train), outline.col = "white", lab = TRUE, title = "Mixed data-type correlation plot for patients without CVD.")
```

# Random Forest Variable Importance Plot
```{r}
rf <- randomForest(Survived~., data = train, ntree=1000, keep.forest=FALSE, importance=TRUE, na.action = na.omit)
varImpPlot(rf)
```

# Feature Selection
```{r}
train <- train[
  c(
    1, #PassengerID
    2, #Pclass
    4, #Sex
    3, #Age
    5, #SibSp
    #6, #Parch
    7, #Fare
    8, #Embarked
    9, #title
    10, #numTicket
    11, #char1Ticket
    12, #charCabin
    13, #meanCabins
    14, #nCabins
    15 #Survived
  )
]
```

# Dummy Vars in Caret
```{r}
dummy <- dummyVars("~.", data=train)
train2 <- data.frame(predict(dummy, newdata = train))
train2
levels(train2$Survived)
```

# Replace Nas with zeroes
```{r}
train2[is.na(train2)] <- 0
train2 <- train2[-(ncol(train2)-1)]
colnames(train2) <- c(colnames(train2[-ncol(train2)]), "Survived")
train2$Survived <- as.factor(train2$Survived)
levels(train2$Survived) <- c("NO", "YES")
```



# Logistic Regression Model
```{r}
default_log_mod = caret::train(
  form = Survived ~ .,
  data = train2,
  trControl = trainControl(method = "repeatedcv", number = 5, classProbs = TRUE, summaryFunction = twoClassSummary, repeats=100),
  method = "glm",
  family = "binomial",
  na.action = "na.exclude",
  metric = "ROC"
)

default_log_mod$results
CV_ConfusionMatrix <- confusionMatrix(default_log_mod)
CV_ConfusionMatrix
```




